---
title: "lecture6-ggplot1"
format: html
---

## **Getting started with data visualization**

Always plot your data!

Two main goals for statistical graphics:

-   To facilitate comparisons

-   To identify trends

There are several ways to make plots in R. We will exclusively use the `ggplot2` package here. If you want to know more about why we’re using `ggplot2` instead of base plot, check out the slides 21+22 from [Jenny Bryan’s ggplot tutorial](https://github.com/jennybc/ggplot2-tutorial/tree/master/ggplot2-tutorial-slides)

## **Load packages**

Now we are ready to begin. `ggplot2` is a graphics package within the `tidyverse`. You should already have this installed, but we need to load the `tidyverse` set of packages before you can use them.

```{r,eval=FALSE}
library(tidyverse)

# If you get an error message, run `install.packages("tidyverse")`  
# If you get an error message talking about the `backports` package, run this first `install.packages("backports")`
```

## **Preparing to create our first `ggplot2` graph**

We will visualize some fuel economy data for 38 popular models of car using the dataset `mpg` which is included in the `ggplot2` package.

```{r,eval=FALSE}
cars 
print(cars) #esto te permite ver la tabla
head(cars,6) #esto permite que veas las primeras 6 filas (puedo cambiar el numero a lo que elija)
tail(cars) #esto es para ver el final de la lista
view(cars) #esto te abre una pestaña con la tabla completa (can´t be seen in a quarto document)
```

